---
title: "Homework 2 – Chapter 3"
subtitle: "Linear regression model using step-wise subsets selection methods"
author: "Akbar Akbari Esfahani"
date: "4/30/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
options(
  "scipen" = 100, 
  "digits" = 4
)
library(MASS)
library(tidyverse)
library(tidymodels)
library(MCMCpack)
```

## Data

The data is from a city assessor that was interested in predicting home sale prices as a function of various characteristics of the home and the surrounding property. The data consists of sale price, finished square feet, # of bedrooms, # of bathrooms, AC, garage size (# of cars fitted), pool, year built, quality (1=high q, 2=med q, 3=low q), style, lot size, and adjacency to highway (1=yes, 0=no). 

> For homework purposes, data processing is not expected 

Read in data and look at variables and data structure
```{r data_read}
re_data <- read_table("data/RealEstateSales.txt") %>% 
  glimpse() %>%
  mutate(
    bed = as.factor(bed),
    bath = as.factor(bath),
    ac = as.factor(ac),
    grsize = as.factor(ifelse(grsize < 3, "2 or less", "3 or more")),
    pool = as.factor(pool),
    age = 2022 - age,
    qu = as.factor(qu)
  )
```

As a first step of performing the required tasks of the homework, we need to create a training and testing set for the data

```{r train_test}
set.seed(314)
# Create a split object
re_split <- 
  initial_split(
    re_data, prop = 0.75, strata = price
  )
# Build training data set
re_training <- 
  re_split %>% 
  training()

# Build testing data set
re_test <- 
  re_split %>% 
  testing()
```

Running the model on all the data to create the base score

```{r modeling}
base_model <- lm(price ~ ., re_training)
summary(base_model)
```
## Backward selection method

We will use the stepAIC function from MASS library

```{r backw_model}
bk_selection <- stepAIC(base_model, direction = "backward")
bk_selection$anova

```

Based on the outcome, the lowest AIC represents the ideal model, generates the following formula:

> price = area + bed + ac + grsize + age + qu + style + lotsize + hgway

with an AIC of 8603.97


```{r bw_fin_mod}
bw_model <- 
  lm(
    price ~ area + bed + ac + grsize + age + qu + style + lotsize + hgway, 
    re_training
  )

pred_test <- 
  predict.lm(
    bw_model,
    re_test,
    interval="prediction"
  ) %>%
  as.data.frame() %>%
  bind_cols(re_test) %>%
  mutate(
    pred_diff = fit - price,
    mean_diff = mean(pred_diff),
    mean_lw_bound = mean(lwr),
    mean_up_bound = mean(upr),
    error_rate = mean_diff/mean(price)
  ) %>%
  dplyr::select(
    mean_diff, mean_lw_bound, mean_up_bound, error_rate
  ) %>%
  distinct()
```

Thus the calculation of the final model from backward selection on the hold out testset yields an error rate of **`r pred_test$error_rate`** with an average prediction interval for price between **`r pred_test$mean_lw_bound` and `r pred_test$mean_up_bound`** while the average predicted price is **`r pred_test$mean_diff`**

## Forward selection method

```{r fw_model}
fw_selection <- stepAIC(base_model, direction = "forward", steps = 10000)
fw_selection$anova
```

Based on the outcome, the lowest AIC represents the ideal model, generates the following formula:

> price = area + bed + bath + ac + grsize + pool + age + qu + style + lotsize + hgway

with an AIC of 8607

```{r fw_fin_mod}
fw_model <- 
  lm(
    price ~ area + bed + bath + ac + grsize + pool + age + qu + style + lotsize + hgway, 
    re_training
  )

pred_test <- 
  predict.lm(
    fw_model,
    re_test,
    interval="prediction"
  ) %>%
  as.data.frame() %>%
  bind_cols(re_test) %>%
  mutate(
    pred_diff = fit - price,
    mean_diff = mean(pred_diff),
    mean_lw_bound = mean(lwr),
    mean_up_bound = mean(upr),
    error_rate = mean_diff/mean(price)
  ) %>%
  dplyr::select(
    mean_diff, mean_lw_bound, mean_up_bound, error_rate
  ) %>%
  distinct()
```

Thus the calculation of the final model from backward selection on the hold out testset yields an error rate of **`r pred_test$error_rate`** with an average prediction interval for price between **`r pred_test$mean_lw_bound` and `r pred_test$mean_up_bound`** while the average predicted price is **`r pred_test$mean_diff`**

## Best subset selection method

```{r best_model}
best_selection <- stepAIC(base_model, direction = "both", steps = 10000)
best_selection$anova
```

Based on the outcome, the lowest AIC represents the ideal model, generates the following formula:

> price = area + bed + ac + grsize + age + qu + style + lotsize + hgway

with an AIC of 8604

```{r best_fin_mod}
best_model <- 
  lm(
    price ~ area + bed + ac + grsize + age + qu + style + lotsize + hgway, 
    re_training
  )

pred_test <- 
  predict.lm(
    best_model,
    re_test,
    interval="prediction"
  ) %>%
  as.data.frame() %>%
  bind_cols(re_test) %>%
  mutate(
    pred_diff = fit - price,
    mean_diff = mean(pred_diff),
    mean_lw_bound = mean(lwr),
    mean_up_bound = mean(upr),
    error_rate = mean_diff/mean(price)
  ) %>%
  dplyr::select(
    mean_diff, mean_lw_bound, mean_up_bound, error_rate
  ) %>%
  distinct()
```

Thus the calculation of the final model from backward selection on the hold out testset yields an error rate of **`r pred_test$error_rate`** with an average prediction interval for price between **`r pred_test$mean_lw_bound` and `r pred_test$mean_up_bound`** while the average predicted price is **`r pred_test$mean_diff`**

## Conclusion

The final model is given by best subset selection method:  
  price = -2303242 + 132.6 x area - 8319.8 x bed - 18433.2 x ac + 20575.0 x grsize + 1224.7 x age - 50768.6 qu - 9046.8 x style + 0.7 x lotsize - 36541.6 x hgway  
  
For homework purposes, this model suffices. However, for enhancement, there will be a follow up with a better analysis